Electron and ion acceleration by relativistic shocks: particle-in-cell simulations

### 2018-01-24 papers/Physica2017/Romansky_Physica2017.tex
paper

In the modeling setup the homogeneous plasma flows into the simulation box through its right boundary
and collides with the reflecting superconducting wall at the left boundary thus launching a shock wave. This  is a common way to initialize a shock (e. g. \cite{Sironi2011}, \cite{Iwamoto2017}), because an initialization using the Rankine-Hugoniot conditions does not take into account the microscopic distribution, and a shock, created in such a way, would probably fall apart into several discontinuities.

@@

The simulations are one-dimensional and have the following parameters of the initial flow: Lorentz factor $\gamma = 1.5$, the number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ , the temperature, $T_e = T_p = 5\cdot10^8 \rm{K}$, the magnetic field $B = 10^{-4} \rm{G}$, the full size of the box $L = 2\cdot10^{12} \rm{cm}$, the number of cells $N=2\cdot10^4$. The electron mass $m_e$ is increase by $100$ times to enlarge electron scales of  the plasma, it is common approach in PIC simulations see e.g. {\cite{Sironi2011}} to reduce required computational recources. The full time of simulation is $T = 5000 {\omega_p}^{-1}$, where $\omega_p$ is electron plasma frequency.These values allow one to define the dimensionless parameter magnetization $\sigma = B^2/4\pi\gamma (n_p m_p + n_e m_e) c^2 = 0.003$. Inclination angle $\theta$ is the angle between the flow velocity and the magnetic field. We present results for the particle spectrum in several simulations at various $\theta$. Stable shocks formed in all simulations and the total energy of the system conserved to better than $3\%$, which indicates that the code provides a correct model of the shock waves.

@@

One can see in Figures \ref{protons} and \ref{electrons} the particle spectra for various inclination angles $\theta$. These spectra consist of three parts - the narrow peak of cold but fast moving initial flow, the wide peak of hot downstream, and the non-thermal accelerated component. Initial flow spectra are compared with Maxwell distribution, shifted on the initial momentum, and hot downstream spectra are compared with Maxwell-J{\"u}ttner distribution.  The small-scale oscilations of spectra are produced by numerical errors and should not be taken into account. Figures show that the spectrum of accelerated particles strongly depends on he inclination angle $\theta$. If $\theta$ is less than the critical value, defined by the equation $c\cdot \cos(\theta_{crit})=v_{shock}$, where all values are measured in the upstream rest frame, particles can escape from the front and cross it several times to gain more energy (see, e.g., \cite{Pelletier2017}), if $\theta$ is greater, shock becomes superluminal - particles need to move along the field lines faster than light to escape from the front, and acceleration is suppressed. It is difficult to evaluate the critical angle \textit{a priori}, because the compression relation of the shock which will be created in the simulation is unknown. However the estimate for the case of a strong wave (the compression relation equals $4$) provides $\theta_{crit}=45^{\circ}$ in the downstream frame, which is consistent with the results of our simulations.

Also, one can see that for the angles less then the critical one, concentrations of accelerated particles are higher and extend to higher enegies for larger $\theta$. The spectrum of non-thermal component is about 10 times higher for the protons than for the electrons at the same energy. It is consistent with other PIC simulations of the shock waves {\cite{Sironi2011}}.

>>>>

In the model setup the homogeneous plasma flows into the simulation box through its right boundary
and collides with the reflecting superconducting wall at the left boundary thus launching a shock wave. This  is a common way to initialize a shock (e. g. \cite{Sironi2011}, \cite{Iwamoto2017}), because an initialization using the Rankine-Hugoniot conditions does not take into account the microscopic distribution, and a shock, created in such a way, would probably fall apart into several discontinuities.

@@

The simulations are one-dimensional and have the following parameters of the initial flow: Lorentz factor $\gamma = 1.5$, the number densities $n_e = 10^{-4} \rm{cm}^{-3}$, $n_p = 10^{-4} \rm{cm}^{-3}$ , the temperature, $T_e = T_p = 5\cdot10^8 \rm{K}$, the magnetic field $B = 10^{-4} \rm{G}$, the full size of the box $L = 2\cdot10^{12} \rm{cm}$, the number of cells $N=2\cdot10^4$. The electron mass $m_e$ is increase by $100$ times to enlarge electron scales of  the plasma, it is common approach in PIC simulations see e.g. {\cite{Sironi2011}} to reduce required computational resources. The full time of simulation is $T = 5000 {\omega_p}^{-1}$, where $\omega_p$ is electron plasma frequency.These values allow one to define the dimensionless parameter magnetization $\sigma = B^2/4\pi\gamma (n_p m_p + n_e m_e) c^2 = 0.003$. Inclination angle $\theta$ is the angle between the flow velocity and the magnetic field. We present results for the particle spectrum in several simulations at various $\theta$. Stable shocks formed in all simulations and the total energy of the system conserved to better than $3\%$, which indicates that the code provides a correct model of the shock waves.

@@

One can see in Figures \ref{protons} and \ref{electrons} the particle spectra for various inclination angles $\theta$. Each spectrum consists of three parts: the narrow peak of cold but moving fast initial upstream flow, the wide peak of hot downstream and the non-thermal accelerated component. The initial upstream flow spectrum is compared with the Maxwell distribution, shifted by the initial momentum, while the hot downstream spectrum is compared with the Maxwell-J{\"u}ttner distribution.  The small-scale oscillations of spectra are produced by numerical errors and should not be taken into account. Figures show that the spectrum of accelerated particles strongly depends on he inclination angle $\theta$. If $\theta$ is less than the critical value, defined by the equation $c\cdot \cos(\theta_{crit})=v_{shock}$, where all values are measured in the upstream rest frame, particles can escape from the front and cross it several times to gain more energy (see, e.g., \cite{Pelletier2017}), if $\theta$ is greater, shock becomes superluminal - particles need to move along the field lines faster than light to escape from the front, and acceleration is suppressed. It is difficult to evaluate the critical angle \textit{a priori}, because the compression relation of the shock which will be created in the simulation is unknown. However the estimate for the case of a strong wave (the compression relation equals $4$) provides $\theta_{crit}=45^{\circ}$ in the downstream frame, which is consistent with the results of our simulations.

Also, one can see that for the angles less then the critical one, concentrations of accelerated particles are higher and extend to higher energies for larger $\theta$. The spectrum of non-thermal component is about 10 times higher for the protons than for the electrons at the same energy. It is consistent with other PIC simulations of the shock waves {\cite{Sironi2011}}.
